LaMachine NLP Software Distribution

### 2018-04-24 docs/lamachine_extabstract_clarin2018.tex
paper update

\author{Maarten van Gompel \\
  Centre for Language and Speech Technology (CLST) \\
  Radboud University Nijmegen, the Netherlands \\
  {\tt proycon@anaproy.nl} \\ %\And % if needed: this makes a second column
%  Second Author \\
%  Department (optional)\\
%  University Name without city \\
%  City, Country \\
% {\tt email@domain} \\
% \AND % if needed: this makes a second row
%  Third Author \\
%  Department (optional)\\
%  University of City, Country \\
%  {\tt email@domain} \\\And
%  Fourth Author \\
%  Department (optional)\\
%  University Name without city \\
%  City, Country \\
%  {\tt email@domain} \\
}

@@

With LaMachine we also attempt to harmonise the metadata of all installed software, by converting metadata upstream repositories to
the CodeMeta standard\footnote{\url{https://codemeta.github.io/}} where possible, or encouraging software developers to provide
their own. This in turn enables other tools to do proper service discovery, and provenance logging.

@@

As LaMachine was conceived initially as the primary means of distribution of the software stack developed at CLST,
Radboud University Nijmegen, it includes a lot of our software. A full list goes beyond the scope of this overview, we
will merely mention some CLARIN/CLARIAH-funded tools: ucto (a tokeniser), Frog (an NLP suite for Dutch), FoLiA (Format
for linguistic Annotation, with assorted tools), FLAT (a web-based linguistic Annotation tool), PICCL (an OCR and
post-OCR correction pipeline) and CLAM. We already include some relevant software by Groningen University and VU
Amsterdam. Moreover, LaMachine incorporates a large number of renowned tools by external parties, offering most notably
a mature Python environment\footnote{scipy, numpy, scikit-learn, matplotlib, nltk, spacy, pytorch, keras, gensim,
etc...}, but also offering space for e.g. R and Java.

\section{Future}

The very recent release of version 2, which constituted a full rewrite, has opened up LaMachine to outside contribution.
Contributor documentation has been written, and at this stage, we greatly welcome external participants to join in. The
main objectives for the future are to providing greater \emph{interoperability} between the included tools through
better \emph{higher-level interfaces} for the researcher. We see this as a bottom-up process and have now established a
firm foundation to build upon. LaMachine development presently takes place in the scope of the CLARIAH WP3 Virtual
Research Environment project, which has much higher ambitions in accommodating the researcher and connectivity of
data and services. A major part of our future focus will be on interoperability with the higher-level tooling developed there, or in the
wider CLARIN/CLARIAH commmunity.

>>>>

\author{Maarten van Gompel \\
  Centre for Language and Speech Technology (CLST) \\
  Radboud University Nijmegen, the Netherlands \\
  {\tt proycon@anaproy.nl} \\ %\And % if needed: this makes a second column
  \url{https://proycon.github.io/LaMachine}
%  Second Author \\
%  Department (optional)\\
%  University Name without city \\
%  City, Country \\
% {\tt email@domain} \\
% \AND % if needed: this makes a second row
%  Third Author \\
%  Department (optional)\\
%  University of City, Country \\
%  {\tt email@domain} \\\And
%  Fourth Author \\
%  Department (optional)\\
%  University Name without city \\
%  City, Country \\
%  {\tt email@domain} \\
}

@@

With LaMachine we also attempt to harmonise the metadata of all installed software, by converting metadata upstream repositories to
the CodeMeta standard\footnote{\url{https://codemeta.github.io/}} \cite{codemeta,codemetar} where possible, or encouraging software developers to provide
their own. This in turn enables other tools to do proper service discovery, and provenance logging.

@@

As LaMachine was conceived initially as the primary means of distribution of the software stack developed at CLST,
Radboud University Nijmegen, it includes a lot of our software. A full list goes beyond the scope of this overview, we
will merely mention some CLARIN-NL/CLARIAH-funded tools: ucto (a tokeniser), Frog (an NLP suite for Dutch), FoLiA (Format
for linguistic Annotation, with assorted tools), FLAT (a web-based linguistic Annotation tool), PICCL (an OCR and
post-OCR correction pipeline) and CLAM. We already include some relevant software by Groningen University and VU
Amsterdam as well. Moreover, LaMachine incorporates a large number of renowned tools by external international parties,
offering most notably a mature Python environment\footnote{scipy, numpy, scikit-learn, matplotlib, nltk, spacy, pytorch,
keras, gensim, etc...}, but also offering space for e.g. R and Java.

\section{Conclusion \& Future work}

The very recent release of LaMachine v2, which constituted a full rewrite, has opened up LaMachine to outside
contribution. Contributor documentation has been written, and at this stage, we greatly welcome external participants
to join in. Aside from the incorporation of new relevant software, the main objectives for the future are to provide
greater \emph{interoperability} between the included tools through better \emph{high-level interfaces} for the
researcher. We see this as a bottom-up process and have now established a firm foundation to build upon. Note that such
proposed interfaces, including the current portal application in LaMachine, are always considered separate independent
software projects, which may be deployed by/in/for LaMachine, but also in other contexts. LaMachine remains `just'
a software distribution at heart.

Development of LaMachine presently takes place in the scope of the CLARIAH WP3 Virtual Research Environment (VRE)
project\footnote{https://github.com/meertensinstituut/clariah-wp3-vre}, which has much higher ambitions in accommodating
the researcher and connectivity of data and services, and transcends also those of the CLARIN LRS Switchboard project
\cite{switchboard}. An important part of our future focus will therefore be on interoperability with the higher-level
tools emerging from the VRE efforts, but also with other parts of the CLARIN infrastructure; single-sign on
authentication being a notable example here.

\bibliographystyle{acl}
\bibliography{lamachine}
